---
title: "BXDharvest_LuminexSerumMetabolites"
author: "Suheeta Roy"
date: "3/15/2021"
output: html_document
---

## Luminex serum metabolites
```{r readBXDlumnxmetabolites}
## set figure options)
options(repr.plot.width = 6, repr.plot.height = 6)
## load library
library(tidyverse)
BXDlumnxmetabolites <- "../data/AgingBXD_Metabolites_LuminexData_Dec2020.csv"
BXDlumnxmetab <- read.csv(BXDlumnxmetabolites)
names(BXDlumnxmetab)
nrow(BXDlumnxmetab)
```

## Plotting Serum Metabolite levels on each diet
```{r stripplotlumnx}
mycolors<- c("#E30613","#312783")
stripchart(BXDlumnxmetab$logInsulin ~BXDlumnxmetab$DietCode, vertical=TRUE, method="jitter",pch=19, cex=0.5,col= mycolors, xlab= "Diets", ylab= "Serum Insulin ( log2 pg/mL)")
mediang <- tapply(BXDlumnxmetab$logInsulin, BXDlumnxmetab$DietCode, median,na.rm=T)
loc <- 1:length(mediang)
segments(loc-0.3, mediang, loc+0.3, mediang, col="black", lwd=3)
stripchart(BXDlumnxmetab$HOMA.IR ~BXDlumnxmetab$DietCode, vertical=TRUE, method="jitter",pch=19, cex=0.5,col= mycolors, xlab= "Diets", ylab= "HOMA-IR")
mediang <- tapply(BXDlumnxmetab$HOMA.IR, BXDlumnxmetab$DietCode, median,na.rm=T)
loc <- 1:length(mediang)
segments(loc-0.3, mediang, loc+0.3, mediang, col="black", lwd=3)
stripchart(BXDlumnxmetab$logLeptin ~BXDlumnxmetab$DietCode, vertical=TRUE, method="jitter",pch=19, cex=0.5,col= mycolors, xlab= "Diets", ylab= "Serum Leptin (log2 pg/mL)")
mediang <- tapply(BXDlumnxmetab$logLeptin, BXDlumnxmetab$DietCode, median,na.rm=T)
loc <- 1:length(mediang)
segments(loc-0.3, mediang, loc+0.3, mediang, col="black", lwd=3)
stripchart(BXDlumnxmetab$logCpeptide ~BXDlumnxmetab$DietCode, vertical=TRUE, method="jitter",pch=19, cex=0.5,col= mycolors, xlab= "Diets", ylab= "Serum C-peptide (log2 pg/mL)")
mediang <- tapply(BXDlumnxmetab$logCpeptide, BXDlumnxmetab$DietCode, median,na.rm=T)
loc <- 1:length(mediang)
segments(loc-0.3, mediang, loc+0.3, mediang, col="black", lwd=3)
```

## T tests for serum metabolites (luminex)
```{r ttest}
t.test(logInsulin~DietCode,data=BXDlumnxmetab)
t.test(logLeptin~DietCode, data= BXDlumnxmetab)
t.test(logCpeptide~DietCode, data= BXDlumnxmetab)
t.test(HOMA.IR~DietCode, data= BXDlumnxmetab)
```

## Looking at correlation between age at death, insulin and diet using linear regression (Chow diet= red, High fat diet= blue)
```{r regressionins}
mycolors<- c("CD"= "#E30613", "HFD"= "#312783")
## Correlation between age at death and serum insulin at harvest
out5<- lm(AgeAtDeath~logInsulin + DietCode, data= BXDlumnxmetab)
summary(out5)
out5a<- lm(AgeAtDeath~logInsulin, subset= BXDlumnxmetab$DietCode== 'CD', data= BXDlumnxmetab)
summary(out5a)
out5b<- lm(AgeAtDeath~logInsulin, subset= BXDlumnxmetab$DietCode== 'HF', data= BXDlumnxmetab)
summary(out5b)
plot( BXDlumnxmetab$AgeAtDeath,BXDlumnxmetab$logInsulin, ylab= "Serum Insulin log2 (pg/mL)", xlab= "Age At Death (Days)",pch=21, cex=0.8, col= "white", bg= mycolors[BXDlumnxmetab$DietCode])
```

## Looking at correlation between age at death, HOMA-IR and diet using linear regression (Chow diet= red, High fat diet= blue)
```{r regressionhoma}
mycolors<- c("CD"= "#E30613", "HFD"= "#312783")
## Correlation between age at death and HOMA-IR at harvest
out6<- lm(AgeAtDeath~HOMA.IR + DietCode, data= BXDlumnxmetab)
summary(out6)
out6a<- lm(AgeAtDeath~HOMA.IR, subset= BXDlumnxmetab$DietCode== 'CD', data= BXDlumnxmetab)
summary(out6a)
out6b<- lm(AgeAtDeath~HOMA.IR, subset= BXDlumnxmetab$DietCode== 'HF', data= BXDlumnxmetab)
summary(out6b)
plot( BXDlumnxmetab$AgeAtDeath,BXDlumnxmetab$HOMA.IR, ylab= "HOMA-IR", xlab= "Age At Death (Days)",pch=21, cex=0.8, col= "white", bg= mycolors[BXDlumnxmetab$DietCode])
```

## Looking at correlation between age at death, leptin and diet using linear regression (Chow diet= red, High fat diet= blue)
```{r regressionlep}
mycolors<- c("CD"= "#E30613", "HFD"= "#312783")
## Correlation between age at death and serum leptin at harvest
out7<- lm(AgeAtDeath~logLeptin + DietCode, data= BXDlumnxmetab)
summary(out7)
out7a<- lm(AgeAtDeath~logLeptin, subset= BXDlumnxmetab$DietCode== 'CD', data= BXDlumnxmetab)
summary(out7a)
out7b<- lm(AgeAtDeath~logLeptin, subset= BXDlumnxmetab$DietCode== 'HF', data= BXDlumnxmetab)
summary(out7b)
plot(BXDlumnxmetab$AgeAtDeath, BXDlumnxmetab$logLeptin, ylab= "Serum Leptin log2 (pg/mL)", xlab= "Age At Death (Days)",pch=21, cex=0.8, col= "white", bg= mycolors[BXDlumnxmetab$DietCode])
```

## Looking at correlation between age at death, c-peptide2 and diet using linear regression (Chow diet= red, High fat diet= blue)
```{r regressioncpep}
mycolors<- c("CD"= "#E30613", "HFD"= "#312783")
## Correlation between age at death and serum c-peptide2 at harvest
out8<- lm(AgeAtDeath~logCpeptide + DietCode, data= BXDlumnxmetab)
summary(out8)
out8a<- lm(AgeAtDeath~logCpeptide, subset= BXDlumnxmetab$DietCode== 'CD', data= BXDlumnxmetab)
summary(out8a)
out8b<- lm(AgeAtDeath~logCpeptide, subset= BXDlumnxmetab$DietCode== 'HF', data= BXDlumnxmetab)
summary(out8b)
plot(BXDlumnxmetab$AgeAtDeath, BXDlumnxmetab$logCpeptide, ylab= "Serum C-peptide log2 (pg/mL)", xlab= "Age At Death (Days)",pch=21, cex=0.8, col= "white", bg= mycolors[BXDlumnxmetab$DietCode])
```
