---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---

```{r}
#install.packages("lme4")
#install.packages("tidyverse")
#install.packages("vioplot")
#install.packages("metafor")
library(tidyverse)
library(vioplot)
library(lme4)
library(metafor)
```

```{r}
BXDagingbwt <- "../data/AgingBXDNOV2018dataSR2.csv"
agingbxd <- read.csv(BXDagingbwt)
names(agingbxd)
```

```{r}
summarizeByDiet <- function(data,pheno,nmin=3)
    {
    strainByDiet <- group_by(data,DietCode,StrainNameCurrent) %>%
          summarize(n=n()) %>%
          pivot_wider(id_cols = "StrainNameCurrent",names_from = "DietCode",values_from = "n")
    strainsWithnmin <- filter(strainByDiet,CD>nmin) %>% filter(HF>nmin) %>% select(StrainNameCurrent)
    newdata <- filter(data,StrainNameCurrent %in% strainsWithnmin$StrainNameCurrent) %>%
             mutate(StrainNameCurrent=fct_drop(fct_explicit_na((StrainNameCurrent))))
    
    # lm
    outLM <- lm(formula(paste(pheno,"~DietCode+StrainNameCurrent")),data=newdata)
    # lmx
    outLMX <- lm(formula(paste(pheno,"~DietCode*StrainNameCurrent")),data=newdata)
    # lmm
    outLMM <- lmer(formula(paste(pheno,"~DietCode+(1+DietCode|StrainNameCurrent)")),data=newdata)
    summary(outLMM)
    # meta analysis
    pheno <- as.name(pheno)
    strainByDiet <- group_by(newdata,DietCode,StrainNameCurrent) %>% 
                summarize(n:=sum(!is.na( {{ pheno }}  )),
                          mean:=mean( {{ pheno }},na.rm=T),
                          sd := sd( {{ pheno }},na.rm=T)) %>%
                mutate(se=sd/sqrt(n)) %>%
                pivot_wider(id_cols = "StrainNameCurrent",
                            names_from = c("DietCode"),values_from = c("n","mean","sd"))

   outMeta <- rma.uni(measure="MD",m1i=mean_HF,m2i=mean_CD,
                      sd1i=sd_HF,sd2i=sd_CD,n1i=n_HF,n2i=n_CD,
                      data=strainByDiet)
    
    lmtab <- as_tibble(coef(summary(outLM)),rownames="coefname") %>% filter(!grepl("Strain",coefname))
    av <- anova(outLM)
    avpct <- round(1000*av[,2]/sum(av[,2]))/10
    names(avpct) <- rownames(av)
    lmxtab <- as_tibble(coef(summary(outLMX)),rownames="coefname") %>%    filter(!grepl("Strain",coefname))
    avx <- anova(outLMX)
    avxpct <- round(1000*avx[,2]/sum(avx[,2]))/10
    names(avxpct) <- rownames(avx)
    lmmtab <- coef(summary(outLMM))
    metatab <- coef(summary(outMeta))
    rownames(metatab) <- "DietCodeHF"
    
    cat("\n --- ANOVA --- \n")
    print(avpct)
    cat("\n --- LM --- \n")
    print(lmtab)
    cat("\n --- ANOVAX --- \n")
    print(avxpct)
    cat("\n --- LMX --- \n")
    print(lmxtab)
    cat("\n --- LMM --- \n")
    print(lmmtab)
    cat("\n --- Meta --- \n")
    print(metatab)

   list(outLM=outLM,outLMX=outLMX,outLMM=outLMM,outMeta=outMeta)
}

out <- summarizeByDiet(agingbxd,"AgeAtDeath")
```

```{r}
summarizeByCovariate <- function(data,pheno,cov,nmin=3)
    {
    covname <- as.name(cov)

    byStrain <- group_by(data,StrainNameCurrent) %>%
          summarize(n=n())
    strainsWithnmin <- filter(byStrain,n>nmin) %>% select(StrainNameCurrent)
    newdata <- filter(data,StrainNameCurrent %in% strainsWithnmin$StrainNameCurrent) %>%
             mutate(StrainNameCurrent=fct_drop(fct_explicit_na((StrainNameCurrent))))
    
    # lm
    outLM <- lm(formula(paste(pheno,"~",cov)),data=newdata)
    # lmm
    outLMM <- lmer(formula(paste(pheno,"~",cov,"+(1|StrainNameCurrent)")),data=newdata)
    # meta analysis
    covLM <- function(df)
        {
        lm(formula(paste(pheno,"~",cov)),data=df)
        }
    pheno <- as.name(pheno)
    strainByCov <- group_by(newdata,StrainNameCurrent) %>% 
                   nest() %>% 
                   mutate(lmresult=map(data,covLM)) %>%
                   mutate(beta=map(lmresult,broom::tidy)) %>% 
                   unnest(beta) %>%
                   filter(term==cov)
    outMeta <- rma.uni(yi=strainByCov$estimate,sei=strainByCov$std.error)

    lmtab <- as_tibble(coef(summary(outLM)),rownames="coefname") %>% filter(!grepl("Strain",coefname))
    av <- anova(outLM)
    avpct <- round(1000*av[,2]/sum(av[,2]))/10
    names(avpct) <- rownames(av)
    lmmtab <- coef(summary(outLMM))
    metatab <- coef(summary(outMeta))
    rownames(metatab) <- cov
    
    cat("\n --- ANOVA --- \n")
    print(avpct)
    cat("\n --- LM --- \n")
    print(lmtab)
    cat("\n --- LMM --- \n")
    print(lmmtab)
    cat("\n --- Meta --- \n")
    print(metatab)

   
   list(outLM=outLM,outLMM=outLMM,outMeta=outMeta)
}
out <- summarizeByCovariate(agingbxd,"AgeAtDeath","Weight02")
```

```{r}
summarizeByCovariate2 <- function(data,pheno,cov,backCov,nmin=3)
    {
    covname <- as.name(cov)

    byStrain <- group_by(data,StrainNameCurrent) %>%
          summarize(n=n())
    strainsWithnmin <- filter(byStrain,n>nmin) %>% select(StrainNameCurrent)
    newdata <- filter(data,StrainNameCurrent %in% strainsWithnmin$StrainNameCurrent) %>%
             mutate(StrainNameCurrent=fct_drop(fct_explicit_na((StrainNameCurrent))))
    
    # lm
    outLM <- lm(formula(paste(pheno,"~",cov,"+",backCov,"+StrainNameCurrent")),data=newdata)
    # lmm
    outLMM <- lmer(formula(paste(pheno,"~",cov,"+",backCov,"+(1|StrainNameCurrent)")),data=newdata)
    # meta analysis
    covLM <- function(df)
        {
        lm(formula(paste(pheno,"~",cov,"+",backCov)),data=df)
        }
    pheno <- as.name(pheno)
    strainByCov <- group_by(newdata,StrainNameCurrent) %>% 
                   nest() %>% 
                   mutate(lmresult=map(data,covLM)) %>%
                   mutate(beta=map(lmresult,broom::tidy)) %>% 
                   unnest(beta) %>%
                   filter(grepl(cov,term))
    outMeta <- rma.uni(yi=strainByCov$estimate,sei=strainByCov$std.error)
                   

   # outMeta <- rma.uni(measure="MD",m1i=mean_HF,m2i=mean_CD,
   #                   sd1i=sd_HF,sd2i=sd_CD,n1i=n_HF,n2i=n_CD,
   #                   data=strainByDiet)
    lmtab <- as_tibble(coef(summary(outLM)),rownames="coefname") %>% filter(!grepl("Strain",coefname))
    av <- anova(outLM)
    avpct <- round(1000*av[,2]/sum(av[,2]))/10
    names(avpct) <- rownames(av)
    lmmtab <- coef(summary(outLMM))
    metatab <- coef(summary(outMeta))
    rownames(metatab) <- cov
    
    cat("\n --- ANOVA --- \n")
    print(avpct)
    cat("\n --- LM --- \n")
    print(lmtab)
    cat("\n --- LMM --- \n")
    print(lmmtab)
    cat("\n --- Meta --- \n")
    print(metatab)
   list(outLM=outLM,outLMM=outLMM,outMeta=outMeta)
}
out <- summarizeByCovariate2(agingbxd,"AgeAtDeath","Weight06","DietCode",20)

```

## Looking at the effect of diet on longevity across all strains and adjusting for strain effect
```{r dieteffect}
out1 <- summarizeByDiet(agingbxd,"AgeAtDeath")
summary(out1$outMeta)
```

## Age at entry into colony and start of HFD is not associated with longevity adjusting for strain effect
```{r entryage}
out2 <- summarizeByCovariate(agingbxd,"AgeAtDeath","EntryAgeInColony")
```

## Looking at the effect of initial weight at entry into aging colony on longevity adjusting for age at entry into colony
```{r initialweighteffect}
out3 <- summarizeByCovariate2(agingbxd,"AgeAtDeath","Weight00Baseline","EntryAgeInColony",4)
summary(out3$outMeta)
```

## Looking at the effect of body weight at 100 days on both diets on longevity adjusting for strain effect
```{r earlyweighteffect}
out4 <- summarizeByCovariate(agingbxd,"AgeAtDeath","Weight02")
summary(out4$outMeta)
```

## Looking at the effect of body weight gain on longevity at 100 days on HFD 
```{r earlyHFDwtgain effect}
hfd <- subset(agingbxd, (agingbxd)$DietCode=='HF')
out5 <- summarizeByCovariate(hfd,"AgeAtDeath","WeightGainPost100days",10)
summary(out5$outLM)
```

## Looking at the effect of body weight on longevity at 400 days on CD 
```{r lateCDwt effect}
cd <- subset(agingbxd, (agingbxd)$DietCode=='CD')
out6 <- summarizeByCovariate(cd,"AgeAtDeath","Weight06", 10)
summary(out6$outMeta)
```

## Looking at the effect of body weight on longevity at 400 days on HFD 
```{r lateHFDwt effect}
hfd <- subset(agingbxd, (agingbxd)$DietCode=='HF')
out7 <- summarizeByCovariate(hfd,"AgeAtDeath","Weight06", 11)
summary(out7$outMeta)
```

## Looking at the effect of body weight gain on longevity at 400 days on HFD
```{r lateHFDwtgain effect}
hfd <- subset(agingbxd, (agingbxd)$DietCode=='HF')
out8 <- summarizeByCovariate(hfd,"AgeAtDeath","WeightGainPost400days",11)
summary(out8$outMeta)
```

## Looking at the effect of diet on body weight at 400 days adjusting for strain effect
```{r dietwteffect}
out9 <- summarizeByDiet(agingbxd,"Weight06")
summary(out9$outMeta)
```

## Fat mass and Adiposity in harvested BXD females
```{r}
BXDagingfat <- "../data/AgingBXDharvest_FatMass.csv"
agingbxdfat <- read.csv(BXDagingfat)
names(agingbxdfat)
```

## Looking at the effect of adiposity in harvested females on longevity adjusting for diet and strain effect
```{r fateffect}
out10 <- summarizeByCovariate2(agingbxdfat,"AgeAtDeath","Total.Adiposity","DietCode", 10)
summary(out10$outMeta)
out11 <- summarizeByCovariate2(agingbxdfat,"AgeAtDeath","Visceral.Adiposity","DietCode", 10)
summary(out11$outMeta)
out12 <- summarizeByCovariate2(agingbxdfat,"AgeAtDeath","Subq.Adiposity","DietCode", 10)
summary(out12$outMeta)
```

## Blood metabolites in harvested BXD females
```{r}
BXDagingmet <- "../data/AgingBXD_harvestedfemales_BloodMetabolites_SR.csv"
agingbxdmetab <- read.csv(BXDagingmet)
names(agingbxdmetab)
```

## Looking at the effect of blood metabolites in harvested females on longevity adjusting for diet and strain effect
```{r metabeffect}
out13 <- summarizeByCovariate2(agingbxdmetab,"AgeAtDeath","Glucose","DietCode", 10)
summary(out13$outMeta)
out14 <- summarizeByCovariate2(agingbxdmetab,"AgeAtDeath","TotalChol","DietCode", 10)
summary(out14$outMeta)
out15 <- summarizeByCovariate2(agingbxdmetab,"AgeAtDeath","Triglycerides","DietCode", 10)
summary(out15$outMeta)
out16 <- summarizeByCovariate2(agingbxdmetab,"AgeAtDeath","FFA","DietCode", 10)
summary(out16$outMeta)
```

## Looking at the effect of blood metabolites in harvested females on body weight adjusting for diet and strain effect
```{r metabeffect}
out13a <- summarizeByCovariate2(agingbxdmetab,"WeightBody","Glucose","DietCode", 10)
summary(out13a$outMeta)
out14a <- summarizeByCovariate2(agingbxdmetab,"WeightBody","TotalChol","DietCode", 10)
summary(out14a$outMeta)
out15a <- summarizeByCovariate2(agingbxdmetab,"WeightBody","Triglycerides","DietCode", 10)
summary(out15a$outMeta)
out16a <- summarizeByCovariate2(agingbxdmetab,"WeightBody","FFA","DietCode", 10)
summary(out16a$outMeta)
```
